---
title: "PropertyRollingSaleData"
author: "Vivek Bejugama, Brandon Lawrence, Jack Imburgia"
date: "October 16, 2016"
output: 
  html_document:
    keep_md: true
---

### Location of the data
The data can be found at https://www1.nyc.gov/site/finance/taxes/property-rolling-sales-data.page.  Download the Excel file and save it to CSV format.

## Data Clean-up

### Install the plyr packages and load in to R.
```{r}
library(plyr)
```


### Read the CSV file.
```{r}
# Change this path to your working directory
setwd("C:\\SASUniversityEdition\\myfolders")
bk <- read.csv("rollingsales_bronx.csv",skip=4,header=TRUE)
```

### Do a prelimary check of the data for missing values or outliers.
```{r}
head(bk)
summary(bk)
str(bk)
```

### Clean/format the data with regular expressions. The pattern "[^[:digit:]]" refers to members of the variable name that start with digits. We use the gsub command to replace them with a blank space. We create a new variable that is a "clean' version of sale.price. And sale.price.n is numeric, not a factor.
```{r}
bk$SALE.PRICE.N <- as.numeric(gsub("[^[:digit:]]","", bk$SALE.PRICE))
count(is.na(bk$SALE.PRICE.N))
```

### Make all variable names lower case.
```{r}
names(bk) <- tolower(names(bk)) 
```

### Get rid of leading digits.
```{r}
bk$gross.sqft <- as.numeric(gsub("[^[:digit:]]","", bk$gross.square.feet))
bk$land.sqft <- as.numeric(gsub("[^[:digit:]]","", bk$land.square.feet))
bk$year.built <- as.numeric(as.character(bk$year.built))
```

### Keep only the actual sales.
```{r}
bk.sale <- bk[bk$sale.price.n!=0,]
plot(bk.sale$gross.sqft,bk.sale$sale.price.n)
plot(log10(bk.sale$gross.sqft),log10(bk.sale$sale.price.n))
```

### Let's look at 1-, 2-, and 3-family homes.
```{r}
bk.homes <- bk.sale[which(grepl("FAMILY",bk.sale$building.class.category)),]
dim(bk.homes)
plot(log10(bk.homes$gross.sqft),log10(bk.homes$sale.price.n))
summary(bk.homes[which(bk.homes$sale.price.n<100000),])
```

### Remove outliers that seem like they weren't actual sales.
```{r}
bk.homes$outliers <- (log10(bk.homes$sale.price.n) <=5) + 0
bk.homes <- bk.homes[which(bk.homes$outliers==0),]
plot(log(bk.homes$gross.sqft),log(bk.homes$sale.price.n))
```

