---
title: "PropertyRollingSaleData"
author: "Vivek Bejugama, Brandon Lawrence, Jack Imburgia"
date: "October 16, 2016"
output: html_document
---

## Loading Libraries and Data
```{r}
library(plyr)
setwd("C:\\SASUniversityEdition\\myfolders")
bk <- read.csv("rollingsales_bronx.csv",skip=4,header=TRUE)
```


## Checking the data
```{r}
head(bk)
summary(bk)
str(bk) 
```


# Cleaning and formating the data with regular expressions
### Replacing bad char to blank with gsub.
```{r}
head(bk$SALE.PRICE,20)  #Before Cleaning Sale.Price data
bk$SALE.PRICE.N <- as.numeric(gsub("[^[:digit:]]","", bk$SALE.PRICE))
head(bk$SALE.PRICE.N,20)  #After Cleaning Sale.Price data
count(is.na(bk$SALE.PRICE.N))
```

### Converting all the variable Names to lowercase.
```{r}
names(bk)  # Before Conversion
names(bk) <- tolower(names(bk)) # make all variable names lower case
names(bk)  # After Conversion
```

## Cleaning up leading digits. 
```{r}
head(bk$gross.square.feet,100)    #Before Clean up
head(bk$land.square.feet,100)     #Before Clean up
head(bk$year.built,100)           #Before Numeric Conversion
bk$gross.sqft <- as.numeric(gsub("[^[:digit:]]","", bk$gross.square.feet))
bk$land.sqft <- as.numeric(gsub("[^[:digit:]]","", bk$land.square.feet))
bk$year.built <- as.numeric(as.character(bk$year.built))
head(bk$gross.sqft,100)           #After Clean up
head(bk$land.sqft,100)            #After Clean up
head(bk$year.built,100)           #After Numeric Conversion
```

# Exploring the Data with Histogram
```{r, echo=FALSE}
attach(bk)
hist(sale.price.n) 
detach(bk)
```


## Keeping only the Actual Sales, Removing Observaitons with No Sales.
```{r}
bk.sale <- bk[bk$sale.price.n!=0,]
plot(bk.sale$gross.sqft,bk.sale$sale.price.n)
plot(log10(bk.sale$gross.sqft),log10(bk.sale$sale.price.n))
```


## Looking at 1-, 2-, and 3-family homes
```{r}
bk.homes <- bk.sale[which(grepl("FAMILY",bk.sale$building.class.category)),]
dim(bk.homes)
plot(log10(bk.homes$gross.sqft),log10(bk.homes$sale.price.n))
summary(bk.homes[which(bk.homes$sale.price.n<100000),])
```


## Removing outliers that seem like they aren't actual sales
```{r}
bk.homes$outliers <- (log10(bk.homes$sale.price.n) <=5) + 0
bk.homes <- bk.homes[which(bk.homes$outliers==0),]
plot(log(bk.homes$gross.sqft),log(bk.homes$sale.price.n))
```